Value Ordering, Discrepancies, and Restarts for Subgraph Algorithms

### 2017-12-14 paper/paper.tex
I have faith in James!

\title{Value Ordering, Discrepancies, and Restarts for Subgraph Isomorphism}
\author{Ciaran McCreesh\thanks{This work was supported by the Engineering and Physical Sciences
    Research Council [grant number EP/026842/1]}\\ University of Glasgow, Glasgow, Scotland \\
    ciaran.mccreesh@glasgow.ac.uk}

@@

\begin{abstract}
    Modern subgraph isomorphism solvers use value-ordering heuristics to direct a constraint
    programming style search. These heuristics are far from perfect, and experiments show that using
    a random value-ordering heuristic with restarts beats conventional search with the best known
    tailored heuristic. This motivates the introduction of a slightly-random value-ordering
    heuristic with similar behaviour to the tailored heuristic: when combined with restarts, this
    new search strategy is over a hundred times more effective on satisfiable instances, and does
    not significantly decrease performance on unsatisfiable instances. In contrast,
    discrepancy-based searches are too expensive to be effective in this setting.
\end{abstract}

@@

We could try to introduce tiebreaking mechanisms to further refine this heuristic. However, given
the extreme cost of making an incorrect branching choice early on in search, we believe it is more
fruitful to investigate alternatives to simple backtracking (depth-first) search. This paper
evaluates two such approaches: discrepancy search (which performs poorly), and restarts (which are
extremely beneficial). To explore different parts of the search space when restarting, we introduce
a new value-ordering heuristic, which adds a small amount of randomness to the ``highest degree
first'' heuristic.  The effects of the changes we propose are shown in
\cref{figure:scatter-by-family}: we can make a state-of-the-art algorithm perform much better on a
large number of satisfiable instances, whilst performing worse only rarely on satisfiable instances,
and never on unsatisfiable instances.  Further, these properties are \emph{not} shared by
discrepancy searches (\cref{figure:scatter-dds}) or by random value-ordering with restarts
(\cref{figure:scatter-random}).

@@

@@

@@

>>>>

\title{Value Ordering, Discrepancies, and Restarts for Subgraph Algorithms}
\author{Ciaran McCreesh\thanks{This work was supported by the Engineering and Physical Sciences
    Research Council [grant number EP/026842/1]}\\ University of Glasgow, Glasgow, Scotland \\
    ciaran.mccreesh@glasgow.ac.uk}

@@

\begin{abstract}
    Modern subgraph isomorphism solvers use value-ordering heuristics to direct a constraint
    programming style search. Although these heuristics are far from perfect (particularly early on
    in search), discrepancy-based searches are too expensive to be effective in this setting.
    However, we show that a random value-ordering heuristic with restarts beats conventional search
    with the best-known tailored heuristic. This motivates the introduction of a slightly-random
    value-ordering heuristic with similar behaviour to the tailored heuristic: when combined with
    restarts, this new search strategy is over ?? a hundred times more effective on satisfiable
    instances, and does not significantly decrease performance on unsatisfiable instances. Finally,
    we show that this strategy also improves two maximum common induced subgraph algorithms.
\end{abstract}

@@

We could try to introduce tiebreaking mechanisms to further refine this heuristic. However, given
the extreme cost of making an incorrect branching choice early on in search, we believe it is more
fruitful to investigate alternatives to simple backtracking (depth-first) search. This paper
evaluates two such approaches: discrepancy search (which performs poorly), and restarts (which are
extremely beneficial). To explore different parts of the search space when restarting, we introduce
a new value-ordering heuristic, which adds a small amount of randomness to the ``highest degree
first'' heuristic.  The effects of the changes we propose are shown in
\cref{figure:scatter-by-family}: we can make a state-of-the-art algorithm perform much better on a
large number of satisfiable instances, whilst performing worse only rarely on satisfiable instances,
and never on unsatisfiable instances.  Further, these properties are \emph{not} shared by
discrepancy searches (\cref{figure:scatter-dds}) or by random value-ordering with restarts
(\cref{figure:scatter-random}). Finally, we show that this strategy is also effective in an
optimisation setting, producing clear benefits in two maximum common induced subgraph algorithms.

@@

\medskip

@@

\section{Maximum Common Subgraph Algorithms}

\begin{figure}[p]
    \centering
    \includegraphics*{gen-graph-scatter-mcsplit.pdf}
    \caption{McSplit speedups}
\end{figure}

\begin{figure}[p]
    \centering
    \includegraphics*{gen-graph-scatter-kdown.pdf}
    \caption{k${\downarrow}$ speedups}
\end{figure}

\begin{figure}[p]
    \centering
    \includegraphics*{gen-graph-mcs.pdf}
    \caption{k${\downarrow}$ and McSplit cumulative curves.}
\end{figure}

@@

\section{Extra Plots not for the Final Paper}
